A Decision Procedure for (Co)datatypes in SMT Solvers

### 2014-12-03 tex/codatadp.tex
minor

\renewcommand{\vec}[1]{\boldsymbol{#1}}
\newcommand{\Ec}{\mathsf{E}}
\newcommand{\tEc}{\mathcal{T}(\Ec)}
\newcommand{\tcEc}{\mathcal{T}^\ast(\Ec)}
\newcommand{\rn}[1]{\textsf{\small #1}}
\newcommand{\cvc}{\textsc{cvc}{\small 4}\xspace}
\newcommand{\teq}{\approx}
\newcommand{\tneq}{\not\teq}
\newcommand{\rem}[1]{\textcolor{red}{[#1]}}
\newcommand{\tester}[1]{is\text{-} #1}
\newcommand{\ror}{\quad \parallel \quad}
\newcommand{\tpath}[2]{\mathcal{P}_{ #2 \rightarrow \_ }( #1 )}
\newcommand{\ttpath}[3]{\mathcal{P}_{ #2 \rightarrow #3 }( #1 )}
\newcommand{\ec}[1]{ [ #1 ] }

@@

Datatypes and codatatypes can be characterized axiomatically. They share basic
properties of constructors and selectors. All properties below range over all
$i$, $j$, $\jPrime$, and $k$ within bounds and to all possible splits of the
$n$-ary constructor $\const{C}_{i\negvthinspace j}$'s argument list into $\bar x,
y, \bar z$:
%
\[
\begin{aligned}[t]
\text{Distinctness:}\quad
  & \forall \bar x, \bar y.\;\, \const{C}_{i\negvthinspace j}(\bar x) \tneq \const{C}_{i\negvthinspace \jPrime}(\bar y) \qquad\text{if $j \not= \jPrime$}
  \\
\text{Injectivity:}\quad
  & \forall \bar x, y, \bar z.\;\, \const{C}_{i\negvthinspace j}(\bar x, y, \bar z) \teq \const{C}_{i\negvthinspace j}(\bar x, y', \bar z) \longrightarrow y \teq y'
  \\
\text{Exhaustiveness:}\quad
  &
  \\
\text{Selection:}\quad
  &
\end{aligned}
\]

@@

    * ``no junk, no confusion''

>>>>

\renewcommand{\vec}[1]{\boldsymbol{#1}}
\newcommand{\Ec}{\mathsf{E}}
\newcommand{\tEc}{\mathcal{T}(\Ec)}
\newcommand{\tcEc}{\mathcal{T}^\ast(\Ec)}
\newcommand{\rn}[1]{\textsf{\small #1}}
\newcommand{\cvc}{\textsc{cvc}{\small 4}\xspace}
\newcommand{\teq}{\approx}
\newcommand{\tneq}{\not\teq}
\newcommand{\rem}[1]{\textcolor{red}{[#1]}}
%\newcommand{\is}[1]{is\text{-} #1}
\newcommand{\is}[1]{\const{is#1}}
\newcommand{\ror}{\quad \parallel \quad}
\newcommand{\tpath}[2]{\mathcal{P}_{ #2 \rightarrow \_ }( #1 )}
\newcommand{\ttpath}[3]{\mathcal{P}_{ #2 \rightarrow #3 }( #1 )}
\newcommand{\ec}[1]{ [ #1 ] }

@@

Datatypes and codatatypes can be characterized axiomatically. They share basic
properties of constructors and selectors. All properties below are implicitly
universally quantified and range over all
$i$, $j$, $\jPrime$, and $k$ within bounds:
% and to all possible splits of the
%$n$-ary constructor $\const{C}_{i\negvthinspace j}$'s argument list into $\bar x,
%y, \bar z$:
%
\[
\begin{aligned}[t]
\text{Distinctness:}\quad
  & %\forall \bar x, \bar y.\;\,
    \const{C}_{i\negvthinspace j}(\bar x) \tneq \const{C}_{i\negvthinspace \jPrime}(\bar y) \qquad\text{if $j \not= \jPrime$}
  \\
\text{Injectivity:}\quad
  & %\forall x_1,\ldots,x_{n_{i\negvthinspace j}} y.\;\,
    \const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}}) \teq \const{C}_{i\negvthinspace j}(x_1,\ldots,x_{k-1},y,x_{k+1},\ldots,x_{n_{i\negvthinspace j}}) \longrightarrow x_k \teq y
  \\
\text{Exhaustiveness:}\quad
  & \is{C}_{i1}(x) \mathrel\lor \cdots \mathrel\lor \is{C}_{im_i}(x)
  \\
\text{Selection:}\quad
  & s_{ij}^k(\const{C}_{i\negvthinspace j}(x_1,\ldots,x_{n_{i\negvthinspace j}})) = x_k
\end{aligned}
\]
%
Expressed in the algebraic jargon,
exhaustiveness ensures that ``no junk'' arises, whereas
distinctness and injectivity ensure that ``no confusion'' occur,
whereas

%Sometimes, the requirements on constructors are stipulated as ``no junk, no
%confusion.''

@@
